Figure 3 and figure 3 supplementary material

### 2022-04-29 Figures/Figure3.Rmd
add Figure3.Rmd

---
title: "Figure 2 and figure 2 supplementary material"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(fig.width = 15, fig.height = 17, echo = FALSE, results = "hide", warning = FALSE, message = FALSE)
```

@@

  title = paste0(estimate, " vs CCLE TIL10 Cancer")
  y_lab = "CCLE TIL10 Cancer"
  final_plot[["ccle"]] <- scatterplot_truevsestimate(merged_subset, estimate, "Epidish_CCLE_TIL10_Cancer.cells", title, color,  x_lab, y_lab)

  # Scatterplot BPRNACan
  title = paste0(estimate, " vs BPRNACan Cancer")
  y_lab = "BPRNACan Cancer"
  final_plot[["bprnacan"]] <- scatterplot_truevsestimate(merged_subset, estimate, "Epidish_BPRNACan_cancer", title, color, x_lab, y_lab)

@@

################################################################################
plot_ridgeplot <- function(data){
 data %>% dplyr::select(-c(ParticipantBarcode, Study)) %>% data.matrix() %>% rcorr() -> cor.table
  final_plot <- list()
  for (ct in c("Lymphocytes", "Macrophages", "Neutrophils")){
    # select strings to remove from signature names
    substring_remove <- c(paste("_",ct, sep=""),
      "_Tirosh_sigmatrix_SuppFig_3-b",
      "_scRNA-Seq",
      "Fig2ab-",
      "_PBMCs_scRNAseq_sigmatrix",
      "sigmatrix_",
      "_Fig2cd")

@@

    # ridgeplot perse
    cols <- c("H&E" = "red", "CBSX" = "yellow", "Epidish" = "blue", "Quantiseq" = "pink", "DeconRNASeq" = "green")

@@

  deconv_subset_summary <- ridgeplot_parse(deconv_data = deconv_subset, signatures_dictionary = signature_dict)

@@

  deconv_subset_summary <- ridgeplot_parse(deconv_data = deconv_subset, signatures_dictionary = signature_dict)
  identical(true_subset$ParticipantBarcode, deconv_subset_summary$Samples) %>% stopifnot()

  to_corr <- bind_cols(true_subset, deconv_subset_summary) %>% dplyr::select(-c(Samples))
  ct_corrplots <- list()
   for (ct in c("Lymphocytes", "Macrophages", "Neutrophils")){
    # select strings to remove from signature names
    substring_remove <- c(paste0("_",ct),
      "_Tirosh_sigmatrix_SuppFig_3.b",
      "_scRNA.Seq",
      "Fig2ab.",
      "_PBMCs_scRNAseq_sigmatrix",
      "sigmatrix_",
      "_Fig2cd")

@@

```{r}
read_tsv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/julien/gemdecan/data/all_deconvolutions/all_deconvolutions_all_TCGA.txt") %>%
  mutate_at("Samples", str_remove_all, ".01A*") %>%
  mutate_at("Samples", str_replace_all, "-", ".") %>% 
  group_by(Samples) %>%
  summarise_all(mean) -> tcga_deconv

read_tsv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/facs/Aran_facs.csv") %>%
  mutate_at("Sample_ID", str_remove_all, ".01A*") %>%
  mutate_at("Sample_ID", str_replace_all, "-", ".") -> Aran_all_facs

read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/HE_fractions.csv") %>%
  dplyr::select(-c(...1)) %>%
  mutate_at("ParticipantBarcode", str_remove_all, ".01A*") %>%
  mutate_at("ParticipantBarcode", str_replace_all, "-", ".") %>%
  distinct(ParticipantBarcode, .keep_all= TRUE) -> HE_all_TCGA # for the READ samples TCGA.AG.5 & TCGA.AG.3

@@

```{r}
# Lists of celltypes of each signature
## signature dictionary construction (can use . as regexp!)
## uncoment print of functions to check that you detect what you want.
## non matching signatures/methods (all 0) are removed.
signature_dict <- list()
signature_dict[["BPRNACan"]] <- list(Lymphocytes = c("B", "CD4", "CD8", "NK"),
                    Macrophages = c("Mono", "M0", "M1", "M2"),
                    Neutrophils = c("Neu"))

signature_dict[["BPRNACanProMet"]] <- signature_dict[["BPRNACan"]]
signature_dict[["BPRNACan3DProMet"]] <- signature_dict[["BPRNACan"]]

signature_dict[["CBSX_Fig2ab-NSCLC_PBMCs_scRNAseq_sigmatrix"]] <- list(Lymphocytes = c("T.cells.CD8", "T.cells.CD4","NKT.cells" ,"B.cells", "NK.cells"),
                                                                         Macrophages = c("Monocytes"),
                                                                         Neutrophils = c())

@@

tcga_deconv %>%
  rename_at(vars(matches("^CBSX")), ~ paste0("CBSX_", .)) %>% 
  rename_at(vars(matches("__")), ~ str_replace_all(., "__", "_")) -> tcga_deconv_ridge

@@

ggsave("Fig2_summary_corrplot.pdf", dpi = "print", width = 250, height = 100, units = "mm", scale = 2.5)
```

@@

for (cancer in unique(HE_all_TCGA$Study)){
  ridgeplots_by_cancer[[cancer]] <- wrapper_ridgeplot(deconv_data = tcga_deconv_ridge, true_estimates = HE_all_TCGA, cancer_abreviation = cancer)
  tpurity_by_cancer[[cancer]] <- tryCatch({ wrapper_tpurity(deconv_data = tcga_deconv, true_estimates = Aran_all_facs, estimate = "ABSOLUTE", cancer_abreviation = cancer) # sometimes there arent enough samples of the tumour type
        },
        error = function(cond) {
          message(paste("not enough FACs samples for ", cancer, "no scatterplot will be generated"))
          return(list(ccle = blank, bprna = blank))
        }
  )
  tpurity_by_cancer[[cancer]]$extra = blank

  ridges <- ggarrange(plotlist = ridgeplots_by_cancer[[cancer]], ncol = 3, common.legend = T, legend = "bottom")
  scater <- ggarrange(plotlist = tpurity_by_cancer[[cancer]], ncol = 3, common.legend = T)
  blank  <- ggplot() + theme_void()
  fullplot <- ggarrange(plotlist = list(scater, ridges, blank), nrow = 3, heights = c(0.15, 0.3, 0.45))
  annotate_figure(fullplot, top = text_grob(cancer, color = "black", size = 40))
  pdf_name <- paste("Sup.fig2_", cancer, ".pdf", sep = "")
  ggsave(pdf_name, dpi = "print", width = 210, height = 297, units = "mm", scale = 2.5)
}
system("pdfunite Sup.fig2_*.pdf Sup.fig2.pdf")
```

>>>>

---
title: "Figure 3 and figure 3 supplementary material"
output: html_notebook
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#knitr::opts_chunk$set(fig.width = 15, fig.height = 17, echo = FALSE, results = "hide", warning = FALSE, message = FALSE)
```

@@

  title = paste0(estimate, " vs MethylCBS Cancer")
  y_lab = "MethylCBS Cancer"
  final_plot[["ccle"]] <- scatterplot_truevsestimate(merged_subset, estimate, "EpiDISH_MethylCBS_cancer", title, color,  x_lab, y_lab)

  # Scatterplot BPRNACan
  title = paste0(estimate, " vs BPmetCan Cancer")
  y_lab = "BPmetCan Cancer"
  final_plot[["bprnacan"]] <- scatterplot_truevsestimate(merged_subset, estimate, "EpiDISH_BPmetCan_cancer", title, color, x_lab, y_lab)

@@

################################################################################
plot_ridgeplot <- function(data){
 data %>% dplyr::select(-c(ParticipantBarcode, Study)) %>% data.matrix() %>% rcorr() -> cor.table
  final_plot <- list()
  for (ct in c("Lymphocytes", "Macrophages", "Neutrophils")){
    # select strings to remove from signature names
    # substring_remove <- c(paste("_",ct, sep=""),
    #   "_Tirosh_sigmatrix_SuppFig_3-b",
    #   "_scRNA-Seq",
    #   "Fig2ab-",
    #   "_PBMCs_scRNAseq_sigmatrix",
    #   "sigmatrix_",
    #   "_Fig2cd")
    substring_remove <- c("nothing")

@@

    # ridgeplot perse
    cols <- c("H&E" = "red", "CIBERSORT" = "yellow", "EpiDISH" = "blue")

@@

  #deconv_subset_summary <- ridgeplot_parse(deconv_data = deconv_subset, signatures_dictionary = signature_dict)
  deconv_subset_summary <- dplyr::select(deconv_subset, matches("Samples|Lymphocytes|Macrophages|Neutrophils"))

@@

  #deconv_subset_summary <- ridgeplot_parse(deconv_data = deconv_subset, signatures_dictionary = signature_dict)
  deconv_subset_summary <- dplyr::select(deconv_subset, matches("Samples|Lymphocytes|Macrophages|Neutrophils"))
  identical(true_subset$ParticipantBarcode, deconv_subset_summary$Samples) %>% stopifnot()

  to_corr <- bind_cols(true_subset, deconv_subset_summary) %>% dplyr::select(-c(Samples))
  ct_corrplots <- list()
   for (ct in c("Lymphocytes", "Macrophages", "Neutrophils")){
    # select strings to remove from signature names
    # substring_remove <- c(paste0("_",ct),
    #   "_Tirosh_sigmatrix_SuppFig_3.b",
    #   "_scRNA.Seq",
    #   "Fig2ab.",
    #   "_PBMCs_scRNAseq_sigmatrix",
    #   "sigmatrix_",
    #   "_Fig2cd")
    substring_remove <- c(paste0("_",ct))

@@

```{r}
tcga_deconv <- read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/julien/gemdecan/data/all_deconvolutions/TCGA_EpiDISH_CIBERSORT_BPmetCan_MethylCBS.csv") %>%
  dplyr::select(-Sample) %>%
  dplyr::rename(Samples = ...1) %>%
  rename_with(stringr::str_replace, 
              pattern = "Neutrophiles$", replacement = "Neutrophils", matches("Neutrophiles")) %>%
  mutate_at("Samples", str_remove_all, ".01A*") %>%
  mutate_at("Samples", str_replace_all, "-", ".") %>% 
  group_by(Samples) %>%
  summarise_all(mean)

Aran_all_facs <- read_tsv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/facs/Aran_facs.csv") %>%
  mutate_at("Sample_ID", str_remove_all, ".01A*") %>%
  mutate_at("Sample_ID", str_replace_all, "-", ".")

HE_all_TCGA <- read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/HE_fractions.csv") %>%
  dplyr::select(-c(...1)) %>%
  mutate_at("ParticipantBarcode", str_remove_all, ".01A*") %>%
  mutate_at("ParticipantBarcode", str_replace_all, "-", ".") %>%
  distinct(ParticipantBarcode, .keep_all= TRUE) # for the READ samples TCGA.AG.5 & TCGA.AG.3

@@

```{r}
# Lists of celltypes of each signature ( NO NEED IN THIS CASE)
## signature dictionary construction (can use . as regexp!)
## uncoment print of functions to check that you detect what you want.
## non matching signatures/methods (all 0) are removed.
signature_dict <- list()
signature_dict[["BPmetCan"]] <- list(Lymphocytes = c("_B$", "CD4", "CD8", "NK", "Treg", "Lymphocytes"),
                    Macrophages = c("Mono", "M0", "M1", "M2", "Macrophages"),
                    Neutrophils = c("Neutrophiles"))

signature_dict[["MethylCBS"]] <- list(Lymphocytes = c("_B", "CD4", "CD8", "NK", "Treg", "Lymphocytes"),
                                                                         Macrophages = c("Mono", "Macrophages"),
                                                                         Neutrophils = c())

@@

# tcga_deconv_ridge <- tcga_deconv %>%
#   rename_at(vars(matches("^CBSX")), ~ paste0("CBSX_", .)) %>% 
#   rename_at(vars(matches("__")), ~ str_replace_all(., "__", "_"))
tcga_deconv_ridge <- tcga_deconv

@@

ggsave("Fig3_summary_corrplot.pdf", dpi = "print", width = 250, height = 100, units = "mm", scale = 1)
```

@@

unique_cancers <- dplyr::filter(HE_all_TCGA, ParticipantBarcode %in% tcga_deconv_ridge$Samples) %>% distinct(Study) %>% deframe()
for (cancer in unique_cancers){
  ridgeplots_by_cancer[[cancer]] <- wrapper_ridgeplot(deconv_data = tcga_deconv_ridge, true_estimates = HE_all_TCGA, cancer_abreviation = cancer)
  tpurity_by_cancer[[cancer]] <- tryCatch({ wrapper_tpurity(deconv_data = tcga_deconv, true_estimates = Aran_all_facs, estimate = "ABSOLUTE", cancer_abreviation = cancer) # sometimes there arent enough samples of the tumour type
        },
        error = function(cond) {
          message(paste("not enough FACs samples for ", cancer, "no scatterplot will be generated"))
          return(list(ccle = blank, bprna = blank))
        }
  )
  tpurity_by_cancer[[cancer]]$extra = blank

  ridges <- ggarrange(plotlist = ridgeplots_by_cancer[[cancer]], ncol = 3, common.legend = T, legend = "bottom")
  scater <- ggarrange(plotlist = tpurity_by_cancer[[cancer]], ncol = 3, common.legend = T)
  blank  <- ggplot() + theme_void()
  fullplot <- ggarrange(plotlist = list(scater, ridges, blank), nrow = 3, heights = c(0.15, 0.3, 0.45))
  annotate_figure(fullplot, top = text_grob(cancer, color = "black", size = 40))
  pdf_name <- paste("Sup.fig3_", cancer, ".pdf", sep = "")
  ggsave(pdf_name, dpi = "print", width = 210, height = 297, units = "mm", scale = 2.5)
}
system("pdfunite Sup.fig3_*.pdf Sup.fig3.pdf")
```
